\documentclass[12pt]{article}

%Includes
\usepackage{color}
\usepackage[normalem]{ulem}
%Title
\title{elementOS Design Document}
\author{Corwin McKnight}


\begin{document}
\maketitle
\tableofcontents
\clearpage
\section{Preamble}
This document describes in great detail the inner workings of the elementOS operating system. It currently is not in working condition, is not in any way stable. The document is currently in development, and edits are welcome. The document will have technical english. If you are looking for a general overview, it does not exist. Attempt to read this. If bengal tigers are in your room, have them attempt to read this; I am \emph{curious}. This document describe the intricate workings of both the operating system, and the kernel. Each is divided by the section header Kernel. Why are you still reading the preamble, the introduction is right below:

\section{Introducion}
element\footnote{By element, I mean the OS.} is a project to create an operating system that incorporates modern ideas and algorithms, while still supporting older machines. By doing this it will be very fast on newer systems, while still being fast on slower systems. It is completely backwards compatible up to the 8086 by using selector functions. elementOS is supposed to be a fast, open source operating system with UNIX-like functionality (and userspace). The project will be split into two parts: the Kernel(113 aka.Ununtrium) and the Operating System(elementOS). Each will be discussed in brief in the following paragraphs.

Ununtrium, the kernel that powers elementOS, will be developed in C and C++ for this project. It will be a Hybrid kernel that acts more like a Monolithic. It will also support multiple architectures, such as ARM(Rpi) and possibly CHIP-8\footnote{While intended as a joke, this might actually work with ununtrium-asm ported to CHIP-8/Super CHIP-8}. Ununtrium will support kernel modules to extend the kernel without recompiling, and can possibly replace or add on to kernel services (adding a new sceduler). Ununtrium will also feature a replay journal to create a very bootstrap initramfs on the fly.

For element, it's mostly just a toy OS. It serves no other purpose than being a research experience. If it proves sucessfull, it will undoubltly be continued in a more serious fashion. It in itself is infact a learning experience, already in its \sout{9th} 10th incarnation. It aims to be lightweight, yet very featured. This is a testbed for my crazy ideas basically.
\section{Design}
For element, most of the design lies in the kernel, but from a non-subjective objective view, you could say elementOS act's like a big ball of strung together components that work together like a decantian(?) market. The people who are shopping are the Messages, and the shops are Packages(more on those soon). The Packages recieve the messages, and they get something from the shop, leave, and go to another shop. The packages communicate either with messaging via. IOSys, or with no interaction at all. Only the kernel main function violates this, as it is executing them.

Packages are kernels' way of organising the different groups of functions. Each pacakge is like a shop in the way that it imports goods (data), and it exports marked up goods (processed data). Each package is also self contained.

So, we know Ununtrium will be divided into Packages which act like components, and are called packages. But what about the userspace? The userspace will be a *nix like one, but will be slightly diffrent. It will have system calls available, but each program will be put into a domain, which is like a protection ring but only for system calls. 

So in the root domian (super user, not '/'), all system calls are available. But in the toor domain (opposite of root), mostly all of the system calls are unavailable. Domains can be switched on the fly.
\subsection{Ununtrium}

\subsubsection{Packages}
Packages are kernels' way of organising the different groups of functions. They are analogus to C++ namespaces, and infact the classes written in C++ use namespaces based on their package, and sub-intent. They logically contain different sub-classes. This is quite usefull for orgainzation. The kernel can continue without some Packages, and the inverse is true too: The kernel can run with aditional packages.


\noindent\begin{tabular}{|l|l|}
\hline
    Package & Function\\
\hline
    Memory   & Manages memory, virtual memory, and the like.\\
	IOSys    & Manages all communications between software and hardware\\
	Cerberus & Manages process sceduling. Has a niceness value and modes.\\
             & Also acts as the watchdog timer for the system\\
    Lock     & Manages the locks in the kernel. \\
             & Subpackage of Cerberus\\
	Arch     & Architecture Specific Code\\
	KMod     & Manages modules. Controls runtime permissions, \\
             &their run time, privileges, and when they are run.\\
	
	VFS      & Manages the file system\\
	Devices  & Manages the loading and facilitating of Drivers.\\
    Drivers  & Contains all kernel builtin drivers (like the screen)\\
	\hline
\end{tabular}
\subsubsection{File Hierarchy}

\begin{tabular}{|l|l|}
\hline
Location & Contains \\
\hline
/ & Root of File System \\
/dev & Holds device file handles\\
/Volumes & Holds mounted file systems (besides the root file system)\\
/Applications & Binary program files\\
/sbin & System binary files\\
/usr/bin & User binary files\\
/Library & Holds data and configuration files for system level apps\\
/usr/Library & Holds data and configuration files for user level apps\\
\hline
\end{tabular}
\section{Behavioral Standards}
Every part of element (including 113) must adhere to the following \emph{guidelines}. Can only be broken by putting a comment in the related source file with a good reason why it has been broken.
\begin{itemize}
  \item Must be for benefit of the end user
  \item Must not kill humans
  \item Must have a useful purpose for all clients (if not, then it belongs as a compile time addition)
  \item Does not bloat the kernel
  \item In addition, does not excessively slow the kernel
  \item Does not compromise the security of the user (without their permission (e.g. root))
\end{itemize}

\subsection{API}
The API across all architectures is very general and standerdised. It does not exist yet.
\subsection{Explicitly Defined Flags}
The kernel has a section for flags that define it's properties. It is not dynamic untill the heap is setup, and some are hardcoded (They are RO: ReadOnly). The PropertyStore works like this:
\subsubsection{PropertyStore}
The property store is an array of keys (The name of the property) and the data. They are stored as \verb"property_propstore_t":

\begin{tabular}{ll}
\hline
Property & Purpose \\
\hline
\verb"char* key" & Name\\
\verb"int id" & Internal Id Number\\
\verb"void* data" & Address to malloc()'d data\\
\verb"int data_size" & sizeof(data)\\
\hline
\end{tabular}
\subsubsection{All surely defined items}

\begin{tabular}{lll}
\hline
RO & Name & Purpose \\
\hline
X & KVER & Kernel Version  \\
X & KNAME & Kenrel Codename \\
\hline
\end{tabular}
\subsection{Options}
The kernel has certain options that are set by \verb+make configuration+ that define static compiler variables. They enable, disable, limit, replace, switch, or free diffrent functions of the kernel.
They are passed to the kernel as compile time switches in clang. They might have a data value, but most of the time them just existing is enough:

\begin{tabular}{|l|l|}
\hline
Name & Purpose \\
\hline

\verb+ARCH+ & The current arch to build code for \\
& Format:Super-Model (x86-64-i7) \\
\hline

\verb+NO_PROGRESSBAR+ & Disables progress bars  \\
\hline

\verb+NO_KERNEL_JOURNAL+ & Removes the replay journal from the kernel \\
\hline

\verb+NO_CRYPTO_CEBERUS+ & Disables Cerberus\\
\hline

\verb+NO_VT+ & Forces one terminal (the physical one)\\
\hline

\verb+ENABLE_DEBUG+ & Enables debug messages \\
\hline

\verb+RIGID_FUNCTIONS+ & Forces selector functions to call the code for\\
& ARCH specific code, \\
\hline

\verb+PROPSTORE_RIGID+ & Forces the PropStore to be Read Only\\
\hline

\verb+DEVMODE+ & All Restrictions Lifted, path to key module\\
\hline

\end{tabular}
\subsection{Functions}
The kernel has a specif list of things it can do, it's functions. There are alot of them, some architecture specific, some for devices. Here are some of the main functions:
\subsubsection{Journaled initramfs}
The journaled initramfs is for the benefit of the boot process. The initramfs itself is not journaled, that would be redundant. Because the kernel is updated with each change to it, it would be a good idea to keep a \emph{backup} in kernel, incase of a missing initram. So a journal is kept of either a bootstrap initramfs, or a old-known-to-work-version. If 113 can't find the initramfs on boot, it replays the journal on a ramdisk created by the vfs (that exists on the file volume. This can be disabled for smaller kernels with disable-backup-initramfs=1 passed to make
\subsubsection{VFS}
The vfs is a mount point based tree. The tree starts with the root (/). The root directory is loaded with the partition that holds the /System directory. Usually sda1. A swap partition is mounted at /.swap(a,b,c). In each mount point, there are a collection of inodes, each with a parent and a child id. The file system driver is called each access not from cache.
\subsubsection{Compile-time Architecture Optimisations}
Make does something wonderful. It passes the ARCH variable to clang as a preprocessor directive. Because it does this, the kernel can optimise based on the architecture.

For dynamic functions, they have a \emph{selector} function \verb+(like strlen)+. This, based on the architecture calls on the \emph{real} function \verb+(strlen_i7).+. This makes the function significantly better.

For a function that will be called hundreds (or \textbf{millions}) of times per second, we have a rigid selector. This requires special compile time options (\verb+COMP_RIGIDFUNCTIONS+ which is paired with \verb+ARCH=arch+) that force the compiler (\underline{clang}) to use a specific version of a function. In the case of the example, it would compile for the \emph{i7} \emph{processor}, regardless of the selected architecture. This works in the case of backwards compatibility (setting the arch as a 8086,as a i7 can run 8086 instructions), but sadly not the reverse (a 8086 can't run i7 instructions. This is obvioulsy bad, because as soon as the kernel does a illegal instruction, the kernel crashes, catches itself, then halts.

This get even more bad if a illegal operation occures before the IDT is initialised. So, when executing a function, it will default to 8086 code when either there is no IDT, or the Arch is unknown by using a switch statement/table for all versions, and the default will be the most compatible. This is useful when there is no optimised version of the function.

\subsubsection{Virtual Terminals}
Virtual terminals (as seen by the VFS) are just files that are treated as screens. Residing in \emph{/dev}, they are simply part of the filesystem. But the kernel treats them as a terminal. Each virtual terminal has a size(in bytes), lenght, and width. 
Size is calculated by: 
\[S=l\times w\] 
The v.t. also have a position(represented by the cursor position). It is calculated by:
\[p=80y + x\]
The original x and y are still stored.

The kernel can switch to a terminal and transfer its output onto the screen. It currently runs at 80 by 24 text mode, which the equation \(b=80*24\) shows us, has only \(b=1920\) charaters; which isn't alot. It is presented in a scroll-type window (using pgup and pgdown).
\subsubsection{Watchdog (Cerberus)}
Cerberusd is the watchdog daemon of the system, using an API to access Cerberus. It has an internal list of all system level processes, and if cerberusd sees that any one of them is killed, it will restart it. krestartguard, however is the actual hardware watchdog designation.
\section{Building}
Building and compiling element uses make. It first passes through.
\subsection{Prerequisites}
To make the kernel (with full functionality) there are some prerequisites that are required
\begin{itemize}
	\item make 
	\item nasm (assembler)
	\item clang (compiler)
	\item binutils (build in ./tools)
	\item gcc
	\item flex
	\item bison
	\item libgmp3-dev; libmpfr-dev; libmpc-dev; libmpc-dev
	\item genisoimage 
	\item doxygen
\end{itemize}
\subsubsection{APT}
To install all prerequisites, do a:
TODO
\subsection{Building the cross compiler}
The cross compiler is gcc, compiling for i586-elf. It is built by executing:

cd [Project Directory]

cd tool

./makebinutils.sh

\noindent The i586-elf-gcc binutils should be built now.
\subsection{Building the kernel}
To build the kernel, just execute make kernel.


\end{document}
